---
title: "Titanic"
author: "Aarij Khawaja"
date: '2019-03-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1) Load the data in RStudio

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
titanic_data <- read.csv('titanic_original.csv')
df <- tbl_df(titanic_data)
df
```

2)  Port of embarkation
The embarked column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df1 <- df %>% 
  mutate(embarked = gsub("^$", "S", df$embarked))
#Unique set of values for embarked to make sure it doesn't include an empty string
df1 %>% select(embarked) %>% unique()
```
Here are the unique values for the column 'embarked'. They do not include any blank or empty string

3) Age
Calculate the mean of the Age column and use that value to populate the missing values.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df2 <- df1 %>% mutate(age = ifelse(is.na(df1$age), mean(df1$age, na.rm=TRUE), df1$age))
df2
```

4) Lifeboat
Fill the empty slots in the boat column with a dummy value 'NA'

```{r echo=FALSE, message=FALSE, warning=FALSE}
df3 <- df2 %>% 
  mutate(boat = gsub("^$", "NA", df2$boat))
df3
```

5) Cabin
Create a new column has_cabin_number which has 1 if there is a cabin number, and 0 otherwise.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df4 <- df3 %>% 
  mutate(has_cabin_number = ifelse(grepl("^$", df3$cabin), 0, 1))
df4
```

6) Submit the project

```{r echo=FALSE, message=FALSE, warning=FALSE}
write.csv(df4, "titanic_clean.csv")
```

